\documentclass{report}

\input{../preamble}
\input{../macros}
\input{../letterfonts}

\title{\Huge{21-235 Math Studies Analysis I}}
\author{\huge{Rohan Jain}}
\date{}

\begin{document}

\maketitle
\newpage% or \cleardoublepage
% \pdfbookmark[<level>]{<title>}{<dest>}
\pdfbookmark[section]{\contentsname}{toc}
\tableofcontents

\pagebreak

\chapter{}
\section{Ordered Fields (Review)}

\dfn{Order}{Let $E$ be a set. An \emph{order} on $E$ is a relation $<$ on $E$ such that for all $x, y, z \in E$:
\begin{enumerate}
    \item (Trichotomy) Exactly one of the following holds: $x < y$, $x = y$, or $x > y$.
    \item (Transitivity) If $x < y$ and $y < z$, then $x < z$.
\end{enumerate}}

\ex{Examples of Ordered Sets}{\begin{enumerate}
    \item This definition develops orders on basic number systems: e.g. $\ZZ$, $\QQ$, and $\RR$.
    \item Define $\lesssim$ on $\ZZ$ as follows: We say that $m \lesssim n$ for $m, n \in \ZZ$ if:
\begin{enumerate}
    \item $m$ is even and $n$ is odd
    \item $m,n$ are even and $m < n$
    \item $m,n$ are odd and $m < n$. 
\end{enumerate}
\end{enumerate}}
\noindent Key Concepts: 
\begin{itemize}
    \renewcommand\labelitemi{--}
    \item upper/lower bounds of sets
    \item bounded sets
    \item max/min
    \item supremum/infimum
    \item supremum/infimum property: An ordered set $E$ satisfies such a property if every nonempty set $A \subseteq E$ that's bounded above/below has a supremum/infimum in $E$.
    \item Fact: sup prop $\implies$ inf prop
\end{itemize}

\dfn{Ordered Field}{Let $\FF$ be a field with order $<$. We say that $\FF$ is an \emph{ordered field} provided that:
\begin{enumerate}
    \item For all $x, y, z \in \FF$, if $x < y$, then $x + z < y + z$.
    \item For all $x, y \in \FF$, if $0 < x$ and $0 < y$, then $0 < x \cdot y$.
\end{enumerate}}

\ex{}{$\QQ$ is a field.}
\noindent Facts of any ordered field:
\begin{enumerate}
    \item $0 < 1$
    \item $\nexists x \in \FF$ such that $x^2 = -1$.
\end{enumerate}

\dfn{Ordered Subfield, Homomorphism, Isomorphism}{Let $\FF$ be an ordered field. 
\begin{enumerate}
    \item A set $\mathbb K \subseteq \FF$ is called an \emph{ordered subfield} if $mathbb K$ is an algeraic subfield and $\mathbb K$ is an ordered field equipped with $<$ from $\FF$.
    \item Let $\mathbb G$ be an ordered field and let $f : \mathbb F \to \mathbb G$. We say that $f$ is an \emph{ordered field homomorphism} if it's a field homomorphism and $f(x) < f(y)$ whenever $x < y$.
    \item $f$ is an ordered field isomorphism if $f$ is an ordered field homomorphism and $f$ is bijective.
\end{enumerate}}
\nt{
\begin{enumerate}
    \item If $f : \mathbb F \to \mathbb G$ is an ordered field homomorphism, $f(\mathbb F)$ is an ordered subfield of $\mathbb G$.
    \item OF property $\implies$ $f$ is injective.
    \item $\therefore$ every ordered field homomorphism $f : \mathbb F \to \mathbb G$ is such that $f$ induces a bijection $f: \mathbb F \to f(\mathbb F) \subseteq \mathbb G$.
\end{enumerate}
}

\thm{$\mathbb Q$ is the smallest ordered field. More precisely, if $\FF$ is an ordered field, then there exists a canonical ordered field homomorphism $f : \QQ \to \FF$.}

\noindent Upshot/notation abuse: We identify $f(\QQ) = \QQ$ to view $\QQ \subseteq \FF$. 
In turn, $\NN \subset \ZZ \subset \QQ \subseteq \FF$. 


\section{Types of Ordered Fields}
\dfn{Archimedean, Dedekind complete}{Let $\FF$ be an ordered field.
\begin{enumerate}
    \item We say that $\FF$ is Archimedean if $\forall 0 < x \in \FF$, $\exists n \in \NN$ such that $n > x$.
    \item We say that $\FF$ is Dedekind complete if it satisfies the supremum property.
\end{enumerate}}

\noindent Facts:
\begin{enumerate}
    \item $\QQ$ is Archimedean.
    \item If $\FF$ is Dedekind complete, then $\forall 0 < x \in \FF$ and $\forall 0< n \in \NN$, $\exists !$  $0 < y \in \FF$ such that $y^n = x$.
    \item $\QQ$ is not Dedekind complete. ($\sqrt 2$ is a counterexample.)
\end{enumerate}

\thm{}{Suppose $\FF$ is a Dedekind complete ordered field. Then $\FF$ is Archimedean.}
\begin{proof}
    If not, then $\NN \subset \FF$ is bounded above, and so the supremum property provides $x \in \FF$ such that $x = \sup \NN$. But then $x - 1$ is an upper bound for $\NN$, so there exists $n \in \NN$ such that $x-1 < n$. Hence $x < n + 1$, which contradicts the definition of $x$ as an upper bound. Therefore, $\FF$ is Archimedean.
\end{proof}

\section{Dedekind Completion}
Throughout this section, let $\FF$ be an Archimedean ordered field.

\dfn{Dedekind cut}{We say a set $\mathcal C \subseteq \FF$ is \emph{Dedekind cut} if:
\begin{enumerate}
    \item $\mathcal C \neq \emptyset$ and $\mathcal C \neq \FF$.
    \item If $p \in \mathcal C$ and $q \in \FF$ such that $q < p$, then $q \in \mathcal C$.
    \item If $p \in \mathcal C$, then $\exists r \in \mathcal C$ such that $p < r$.
\end{enumerate}
We will write $\FF^*$ for the set of all Dedekind cuts in $\FF$. It is called the \emph{Dedekind completion} of $\FF$.}

\nt{Let $\mathcal C \subseteq \FF$ be a cut. Then:
\begin{enumerate}
    \item If $p \in \mathcal C$, then $q \notin \mathcal C$, then $p < q$.
    \item If $r \notin \mathcal C$, and $r < s \in \FF$, then $s \notin \mathcal C$.
\end{enumerate}}

\ex{Cut examples}{\begin{enumerate}
    \item Let $q \in \FF$ and define $\mathcal C_q = \{ p \in \FF \mid p < q\}$. Then $\mathcal C_q$ is a cut.
    \begin{proof}
        \begin{enumerate}
            \item $q - 1 < q \implies q - 1 \in \mathcal C_q$. $q \not < q \implies q \notin \mathcal C_q \implies \mathcal C_q \neq \FF$.
            \item Let $p \in \mathcal C_q$. Suppose $s \in \FF$ such that $s < p$. Then $s < q \implies s \in \mathcal C_q$.
            \item Let $p \in \mathcal C_q$. Then $p < \frac{p + q}{2} < q \implies \frac{p + q}{2} \in \mathcal C_q$.
        \end{enumerate}
    \end{proof}
    \item Suppose $\FF$ is such that $\nexists x \in \FF$ such that $x^2 = 2$. Let $\mathcal C = \{ p \in \FF \mid p \leq 0 \text{ or } 0 < p^2 < 2\}$. Then $\mathcal C$ is a cut.
    \begin{proof}
        \begin{enumerate}
            \item $1 \in \mathcal C$ and $1^2 = 1 < 2$. $2 \notin \mathcal C$ and $2^2 = 4 > 2$. 
            \item Let $p \in \mathcal C$ and $q \in \FF$ such that $q < p$. If $q \leq 0$, then $q \in \mathcal C$ trivially. Suppose $0 < q < p$. Then $0 < q^2 < p^2 < 2$, so $q \in \mathcal C$.
            \item Let $p \in \mathcal C$. If $p \leq 0$, then $1 \in \mathcal C$ and $p < 1$, so we're done. Suppose $0 < p^2 < 2$. Note, $0 < 2 - p^2$, so $\frac{2p + 1}{2 - p^2} > 0$. Then we can define $r = 1 + \frac{2p + 1}{2 - p^2} \geq \max(1, \frac{2p + 1}{2 - p^2})$. Then $(p + 1/r)^2 = p^2 + \frac{2p}{r} + \frac{1}{r^2}$. We have:
            \begin{align*}
                p^2 + \frac{2p}{r} + \frac{1}{r^2} &< p^2 + \frac{2p}{r} + \frac 1r \\
                &= p^2 + \frac{2p+1}{r} \\
                &\leq p^2 + 2-p^2 \\
                &= 2.
            \end{align*}
            So, $p < p + 1/r < 2$ and $p + 1/r \in \mathcal C$.
        \end{enumerate}
    \end{proof}
\end{enumerate}}

\subsection{Ordering $\FF^*$}

\mlenma{}{The following hold:
\begin{enumerate}
    \item If $\mathcal A, \mathcal B \in \FF^*$, then exactly one holds:
    \begin{itemize}
        \item $\mathcal A \subset \mathcal B$
        \item $\mathcal A = \mathcal B$
        \item $\mathcal B \subset \mathcal A$
    \end{itemize}
    \item If $\mathcal A, \mathcal B, \mathcal C \in \FF^*$ and $\mathcal A \subset \mathcal B$ and $\mathcal B \subset \mathcal C$, then $\mathcal A \subset \mathcal C$.
\end{enumerate}}
\begin{proof}
    Proof of $2$ is trivial, as well as the equality part for $1$. 
    \begin{itemize}
        \item If $\mathcal A = \mathcal B$, we're done.
        \item Suppose $\exists b \in \mathcal B \setminus\mathcal A$. If $a \in \mathcal A$, then $a < b$, but $\mathcal B$ is a cut so $a \in \mathcal B$, so $\mathcal A \subset \mathcal B$.
        \item Suppose $\exists a \in \mathcal A \setminus \mathcal B$. Then $a < b$ for all $b \in \mathcal B$, so $a \in \mathcal B$, so $\mathcal B \subset \mathcal A$.
    \end{itemize}
\end{proof}


\dfn{Order on cuts}{Given $\mathcal A, \mathcal B \in \FF^*$, we say that $\mathcal A < \mathcal B$ if $\mathcal A \subset \mathcal B$. The lemma above shows that this is infact an order.}


\mlenma{}{Let $E \subseteq \FF^*$ be nonempty and bounded above. Then $\mathcal B = \bigcup_{\mathcal A \in E} \mathcal A$ is a cut.}
\begin{proof}
    \begin{enumerate}
        \item Since $E \neq \emptyset$, there exists $\mathcal A \in E$. So $\mathcal A \neq \emptyset$, hence $\mathcal B \neq \emptyset$. 
        
        Since $E$ is bounded above, there exists $\mathcal C \in \FF^*$ such that $\mathcal A \subset \mathcal C$ for all $\mathcal A \in E$. Since $\mathcal C$ is a cut, there is $q \in \FF$ such that $q \notin \mathcal C$. Then $q \notin \mathcal A$ for all $\mathcal A \in E$, so $q \notin \mathcal B$.
        \item Let $p \in \mathcal B$ and $q \in \FF$ such that $q < p$. Since $\mathcal B$ is a union of cuts, it follows that $p \in \mathcal A$ for some $\mathcal A \in E$. Since $\mathcal A$ is a cut, $q \in \mathcal A \subseteq \mathcal B$.
        \item Let $p \in \mathcal B$. Then $p \in \mathcal A$ for some $\mathcal A \in E$. Since $\mathcal A$ is a cut, there exists $r \in \mathcal A$ such that $p < r$. Since $\mathcal A \subset \mathcal B$, we have $r \in \mathcal B$.
    \end{enumerate}
\end{proof}

\thm{}{$\FF^*$ equipped with the order $<$ satisfies the supremum property.}



\end{document}