\documentclass{report}

\input{../preamble}
\input{../macros}
\input{../letterfonts}
\usepackage{fix-cm}
\usepackage[T1]{fontenc}

\title{\Huge{21-610}\\Algebra I}
\author{\huge{Rohan Jain}}
\date{}

\begin{document}

\maketitle
\newpage% or \cleardoublepage
% \pdfbookmark[<level>]{<title>}{<dest>}
\pdfbookmark[section]{\contentsname}{toc}
\tableofcontents

\pagebreak

\chapter{}
\section{1/17 - Group Actions}

\dfn{Action}{With a group $G$ and set $X$, an \emph{action} of $G$ on $X$ is a HM from $G$ to $\Sigma_X$ (the group of permutations of $X$).}

\dfn{$g \cdot x$}{If $\phi : G \to \Sigma_X$ is an action, then for $g \in G$ and $x \in X$, we write $g \cdot x$ for $\phi(g)(x)$.}
\nt{People will eventually lose the $\cdot$. So, $g \cdot x$ will be written as $gx$.}
\ex{actions}{\begin{itemize}
    \item $1 \cdot x = x$ $(*)$
    \item $g \cdot (h \cdot x) = \phi(g)(\phi(h)(x)) = (\phi(g) \circ \phi(h))(x) = \phi(gh)(x) = (gh) \cdot x$ $(**)$
\end{itemize}}
If $\cdot : G \times X \to X$ satisfies $(*)\cdot \& (**)$, then there's unique action $\phi : G \to \Sigma_X$ such that $g \cdot x = \phi(g)(x)$.
\begin{proof}
    Define $\phi: G \to ^X X$ by $\phi(g)(x) = g \cdot x$. 

    $\phi(g^{-1})$ is 2-sided inverse of $\phi(g)$, $\phi(g) \in \Sigma_X$. So $\phi$ is an HM by $(**)$.
\end{proof}

\dfn{Orbit Equivalence Relation}{Let $G$ act on $X$. The \emph{orbit equivalence relation} on $X$ is induced by action: $x \sim y$ if $\exists g \in G$ such that $g \cdot x = y$.}
\dfn{Orbits}{The equivalence classes of this relation are called \emph{orbits}. They are defined as \[O_x = \{y : x \sim y\} = \{y : \exists g, g\cdot x = y\}\]}

\dfn{Stabilizer}{Let $G$ act on $X$. The \emph{stabilizer} of $x \in X$ is the subgroup of $G$ defined as \[G_x = \{g \in G : g \cdot x = x\}\]}
\newpage
Note that $G_x \leq G$.

\begin{proof}
    We need to show that $G_x$ is a subgroup of $G$.
    \begin{itemize}
        \item $1 \cdot x = x$, so $1 \in G_x$.
        \item $g \in G_x \implies g \cdot x = x \implies g^{-1} \cdot x = x \implies g^{-1} \in G_x$.
        \item $g,h \in G_x$. $(gh) \cdot x = g \cdot (h \cdot x) = g \cdot x = x$, so $gh \in G_x$.
    \end{itemize}
\end{proof}
A calculation:
\begin{align*}
    g_1 \cdot x = g_2 \cdot x &\iff g_2^{-1} \cdot (g_1 \cdot x) = x \\
    &\iff (g_2^{-1}g_1) \cdot x = x \\
    &\iff g_2^{-1}g_1 \in G_x \\
    &\iff g_1 \in g_2G_x \\
    &\iff g_1G_x = g_2G_x
\end{align*}
This gives a bijection between $O_x$ and set of left cosets of $G_x$. So, we have the orbit-stabilizer theorem:
\thm{Orbit-Stabilizer Theorem}{Let $G$ act on $X$. Then for all $x \in X$, $|O_x| = [G : G_x]$.}
\dfn{Fixed Point}{Let $G$ act on $X$. A \emph{fixed point} of the action is an $x \in X$ such that $g \cdot x = x$ for all $g \in G$. That is, $G_x = G$.}
\dfn{Fixed-Point Set}{Let $G$ act on $X$. Choose a $g \in G$. The \emph{fixed-point set} of $g$ is the set of all $x \in X$ such that $g \cdot x = x$ and is denoted $X_g$.}
\section{1/19 - Group Actions}
\ex{Automorphism Groups}{$$\Aut(G) = \{f : G \to G : f \text{ is an isomorphism}\}$$
$\phi \in \Sigma_G$, $\phi(ab) = \phi(a)\phi(b)$. Recall conjugate of $h$ by $g$ is $h^g = ghg^{-1}$.

Fact 1: For any $g \in G$, $h \mapsto h^g$ is an automohism of $G$.

Fact 2: If $\phi : G \to \Aut(G)$, $\phi : g \mapsto (h \mapsto h^g)$, then $\phi$ is an HM for $G$ to $\Aut(G)$.

$G$ acts on $G$ by automorphisms. $g \cdot h = h^g = ghg^{-1}$.

In this setting:
\begin{enumerate}
    \item Orbit equivalence relation is conjugacy.
    \item Orbits are conjugacy classes.
    \item For $h \in G$, the stabilizer of $h$ for conjugation action $= \{g : h^g = g\}$. \begin{align*}
        h^g = h &\iff ghg^{-1} = h \\
        &\iff gh = hg \\
        &\iff g \in C_G(h)
    \end{align*}
\end{enumerate}
}

\dfn{Centralizer}{Let $G$ act on $X$. The \emph{centralizer} of $x \in X$ is the subgroup of $G$ defined as \[C_G(x) = \{g \in G : g \cdot x = x \cdot g\}\]}

\thm{Orbit-Stablizer Equation for conjugation action}{$$|\text{conj class of }h| = [G : C_G(h)]$$}
$$|G| = \sum_{C \text{ conj. class} }|C|$$
So, if $C = $ class of $h$, $|C| = [G : C_G(h)]$.

Recall the definition of a fixed-point. So, for $G$ acting on $G$ by conjugation, $X_g = C_G(g)$. That is,
\begin{align*}
    h \text{ fixed point} \iff h^g = h \iff hg = gh \tag{for all $g$}
\end{align*}
\dfn{Center}{The \emph{center} of $G$ is $Z(G) = \{g \in G : gh = hg \text{ for all } h \in G\}$. In fact, $Z(G)$ is normal in $G$. That is, $Z(G) \lhd G$.}
\thm{}{Let $p$ be prime. Let $G$ be a group of order $p^n$. Then $Z(G) \neq 1$.}
\begin{proof}
    Let $G$ act on $G$ by conjugation. $G$ is partitioned into orbits (i.e. conjugacy classes). 

    For any $h$, we know that the size of the class of $h$ is $[G : C_G(h)]= \dfrac{p^n}{|C_G(h)|}$. Each orbit has size $1$ or a power of $p$. So, $|C_G(h)|$ is a power of $p$.

    Note in any action of $G$ onto $X$, $x$ being a fixed point implies $O_x = \{x\}$. So, $|O_x| = 1$.

    So, $|G| = A + B$ where $A$ is the number of orbits of size 1 and $B = \sum |C|$ where $C$ is a conjugacy class of size $p^n$ for $n > 0$. 

    So, $A = p^n - B$. So $p | A$. As $Z(G) \neq \emptyset$, $|Z(G)| > 0, p | |Z(G)|$. So, $|Z(G)| \geq p$, which is at least 2, so $Z(G) \neq 1$.
\end{proof}
\thm{Cauchy's Theorem}{Let $G$ be a finite group. If $p$ is a prime dividing $|G|$, then $G$ has an element or subgroup of order $p$.}

Facts to remember from undergraduate group theory:
\begin{itemize}
    \item Let $N \lhd G$. Then subgroups of $G/N$ are in bijection with $\{H : N \leq H \leq G\}$. In fact $H \mapsto H/N$ is a bijection.
    \item Normal subgroups of $G/N$ are uniquely of the form $H/N$ where $H \lhd G$ and $N \leq H$.
    \item $H/N \lhd G/N$ , $\frac{G/N}{H/N} \cong G/H$.
\end{itemize}
\newpage
\section{1/22 - Using Group Actions to Prove Theorems}
Now we prove Cauchy's Theorem:
\begin{proof}
    Let $X = \{ (g_1, \ldots, g_p) \in G^p : g_1 \cdots g_p = 1\}$. Some remarks:
    \begin{itemize}
        \item $(g_1, \ldots, g_p) \in X \iff (g_1\ldots g_{p-1})g_p = 1$. So, $g_p = (g_1\ldots g_{p-1})^{-1}$ and $(g_p, g_1, \ldots, g_{p-1}) \in X$. So, $|X| = |G|^{p-1}$.
        \item $X \neq \emptyset$ as $(1, \ldots, 1) \in X$.
    \end{itemize}
    So now it's easy to define an action of $C_p$(cyclic group of order $p$) on $X$. Explicitly, if $C_p = \langle a \rangle>$, then $a \cdot (g_1, \ldots, g_p) = (g_2, \ldots, g_p, g_1)$. 

    Now we analyze the fixed-points. $(g_1, \ldots, g_p)$ if and only if all the $g_i$ are equal. So, fixed points in the action of $C_p$ on $X$ are $(g, \ldots, g) \in X$ where $g^p = 1$. 

    As $p \mid |G|$, $p \mid |X| = |G|^{p-1}$. As $p$ is prime, $|C_p| = p$. So all orbits have size 1 or $p$. $X$ is partitioned into orbits, say $$|X| = C + D_p$$
    So $p | C$ where $C$ is the number of fixed-points for this action. As $(1, \ldots, 1)$ is a fixed-point, $C>0$, so $C \geq p > 1$. So, there is a fixed-point $(g, \ldots, g) \in X$ where $g^p = 1$. So, $g$ has order $p$.
\end{proof}
\dfn{$\Syl_p(G)$}{$\Syl_p(G) = \{H : H \leq G, |H| = p^k \text{ for some } k \geq 1 \text{ for largest } k\}$}
\nt{If $p \nmid |G|$, then $\Syl_p(G) = \{1\}$.}
\thm{Sylow's Theorem}{Let $G$ be a finite group. Let $p$ be a prime dividing $|G|$. \begin{enumerate}
    \item If $H \leq G$ and $|H|$ is a power of $p$, there is $K \in \Syl_p(G)$ such that $H \leq K$.
    \item If $K_1, K_2 \in \Syl_p(G)$, then $K_1$ and $K_2$ are conjugate.
    \item $|\Syl_p(G)| \equiv 1 \pmod{p}$ and divides $|G|$.
\end{enumerate}}
\noindent Notes before proof:
\begin{itemize}
    \item Let $G$ be a group. $\alpha \in \Aut(G), H\leq G$. Then $\alpha [H] = \{\alpha(h) : h \in H\}$ is a subgroup of $G$  and $\alpha[H] \cong H$. $\alpha$ is a bijection from $H$ to $\alpha[H]$. 
    \item In particular, for $g \in G$, if $\alpha$ is ``conjugation by $g$'', then $\alpha[H] = gHg^{-1}$ or $H$ goes to $H^g$. We can check: $G$ acts on $\{H : H \leq G\}$. $g\cdot H = H^g = gHg^{-1}$.
    \item $H$ is a fixed point of this action if and only if $H^g = H$ for all $g \in G$. That is, $H \lhd G$.
    \item For any $H$, stabilizer of $H$ for this action is $N_G(H) = \{g \in G : gHg^{-1} = H\}$.
\end{itemize}
\dfn{Normalizer}{Let $H \leq G$. The \emph{normalizer} of $H$ in $G$ is $N_G(H) = \{g \in G : gHg^{-1} = H\}$.}
\begin{itemize}
    \item Let $G$ act on $X$. Then we know that if $Y \subseteq X$, and $g \cdot y \in Y$ for all $g \in G$ and $y \in Y$, then $Y$ is a union of orbits and we then get an action for $G$ onto $Y$.
    \item Let $G$ act on $X$. Let $H \leq G$, now easily $H$ acts on $X$. Each $G$-orbit breaks up as a union of $H$-orbits.
\end{itemize}
\newpage
\section{1/26 - Using Group Actions to Prove Theorems}
Now we finally prove Sylow's.
\begin{proof}
    
\end{proof}

\section{1/29 - Series in Groups}
\dfn{Subnormal Series}{A \emph{subnormal series} for a group $G$ is a sequence $(G_i)$ of subgroups \[1  = G_0 \lhd G_1 \lhd \cdots \lhd G_n = G\]}
\dfn{Normal Series}{A subnormal series is a \emph{normal series} if $G_i \lhd G$ for all $i$.}
\dfn{Characteristic Series}{A \emph{characteristic series} is a normal series $(G_i)$ such that $G_i \text{char} G$ for all $i$.}
\dfn{Commutator}{Let $G$ be a group. The \emph{commutator} of $g,h \in G$ is $[g,h] = ghg^{-1}h^{-1}$.}
\nt{$[g,h] = 1$ if and only if $gh = hg$.}
\dfn{Commutator Subgroups}{The \emph{commutator subgroup} of $G$ is $[G,G] = \langle [g,h] : g,h \in G\rangle$.}
\nt{If $\alpha \in \Aut(G)$, then $\alpha([g,h]) = [\alpha(g), \alpha(h)]$. So $\alpha([G,G]) = [G,G]$. So $[G,G] \lhd G$.}
\noindent What does it mean for $G/N$ to be abelian? We get:
\begin{align*}
    &[aN, bN] = 1 \forall a ,b \\
    &\iff [a,b]N = 1 \\
    &\iff [a,b] \in N \\
    &\iff [G,G] \leq N
\end{align*}
So $[G, G]$ is the least normal subgroup of $G$ such that $G/N$ is abelian.
\dfn{Solvable}{A group $G$ is \emph{solvable} iff there is a subnormal series $(G_i)$ in $G$ such that $G_{i+1}/G_i$ is abelian for all $i < n$.}
Trivially, abelian groups are solvable. But there are non-abelian solvable groups. For example, $S_3$ is solvable. This is because $1 \lhd A_3 \lhd S_3$. $S_3/A_3 \cong C_2$ is abelian.
\newpage
\thm{}{If $G$ is solvable and $H \leq G$, then $H$ is solvable.}
\begin{proof}
    Let $(G_i)$ be a subnormal series for $G$ such that $[G_{i+1}, G_{i+1}] \leq G_i$ for all $i < n$. Then $(H \cap G_i)$ is a subnormal series for $H$ such that $[H \cap G_{i+1}, H \cap G_{i+1}] \leq H \cap G_i$ for all $i < n$.
\end{proof}

\thm{}{If $G$ is solvable and $N \lhd G$, then $G/N$ is solvable.}
\begin{proof}
    Note: If we have $N \lhd G$ and $H \leq G$ and $\phi_N : G \to G/N$ is the natural homomorphism, then $\phi_N(H) \leq  = HN/N \leq G/N$.

    Let $(G_i)$ be a subnormal series for $G$ such that $[G_{i+1}, G_{i+1}] \leq G_i$ for all $i < n$. Then $(G_iN/N)$ is a subnormal series for $G/N$ such that $[G_{i+1}N/N, G_{i+1}N/N] \leq G_iN/N$ for all $i < n$.
\end{proof}

\section{1/31 - }
\thm{}{If $N \lhd G$ with $N$ and $G/N$ both solvable, then $G$ is solvable.}
\begin{proof}
    Let $(N_i)$ be a subnormal series for $N$ such that $[N_{i+1}, N_{i+1}] \leq N_i$ for all $i < n$. Let $(V_i)$ be a subnormal series for $G/N$ such that $[V_{i+1}, V_{i+1}] \leq V_i$ for all $i < n$.

    For each $j$, let $G_j$ be the unique subgroup of $G$ such that $N \leq G_j$ and $G_j/N = V_j$. 

    What we want is that $N_0, N_1, \ldots, N_m = N = G_0, G_1, \ldots, G_m = G$ is a subnormal series for $G$ such that $[G_{i+1}, G_{i+1}] \leq G_i$ for all $i < n$.

    Since $V_j \lhd V_{j+1}$, run time to check that $G_j \lhd G_{j+1}$. We know that $N_{j+1} / N_j$ and $V_{j+1} / V_j$ are abelian. So we have that $V_{j+1} / V_j = G_{j+1} / N / G_j / N \cong G_{j+1} / G_j$ is abelian. So $G_{j+1} / G_j$ is abelian. So $G$ is solvable.
\end{proof}

\dfn{Derived Series}{Let $G$ be a group. Then the \emph{derived series} of $G$ is a sequence of subgroups of $G$ defined as \[G^{(0)} = G, G^{(1)} = [G,G], G^{(2)} = [G^{(1)}, G^{(1)}], \ldots, G^{(n+1)} = [G^{(n)}, G^{(n)}]\]}

\nt{$G^{(n)} / G^{(n+1)}$ is abelian. In fact, $G^{(n+1)}$ is the best normal subgroup of $G^{(n)}$ such that $G^{(n)} / G^{(n+1)}$ is abelian.}
\nt{$G^{(n)} \text{char} G$ for all $n$.}
\mprop{}{If there is $n$ with $G^{(n)} = 1$, then $G$ is solvable.}
\begin{proof}
    $G^{(n)} = 1 \implies G^{(n-1)}$ is abelian. So $G^{(n-1)} / G^{(n)} = G^{(n-1)}$ is abelian. So $G$ is solvable.
\end{proof}
\newpage
\mprop{}{If $G$ is solvable, then there is $n$ with $G^{(n)} = 1$.}
\begin{proof}
    Let $(G_i)$ be a subnormal series for $G$ such that $[G_{i+1}, G_{i+1}] \leq G_i$ for all $i < n$. We show by induction on $s$ that $G^{(s)} \leq G_{n-s}$ for $0 \leq s \leq n$. 

    If this works, then $G^{(n)} \leq G_0 = 1$. So $G^{(n)} = 1$.

    \textbf{Base Case:} $s = 0$. $G^{(0)} = G \leq G_n = G$.

    Now suppose $s < n$ and $G^{(s)} \leq G_{n-s}$. Then $G^{(s+1)} = [G^{(s)}, G^{(s)}] \leq [G_{n-s}, G_{n-s}] \leq G_{n-s-1}$.

    So $G^{(n)} = 1$.
\end{proof}

\dfn{Simple}{$G$ is \emph{simple} if $G\neq 1$ and $N \lhd G$ implies $N = 1$ or $N = G$.}

\cor{}{If $G$ is abelian and simple, then $G$ is cyclic of prime order.}

\thm{}{For $n \geq 5$, $A_n$ is simple.}
\thm{}{If $G$ has a nonabelian simple subgroup, then $G$ is not solvable.}
\begin{proof}
    \begin{enumerate}[label=(\alph*)]
        \item If $K$ is simple and nonabelian, $K$ is definitely not solvable.
        \item Subgroups of solvable groups are solvable.
    \end{enumerate}
\end{proof}

\section{2/2 - (absent)}
\dfn{Nilpotence}{A group $G$ is \emph{nilpotent} if there is a subnormal series $(G_i)$ for $G$ such that $G_{i+1} \lhd G_i / G_{i+1}$ for all $i$.}

\section{2/5 - Nilpotency}
\dfn{Central Series}{A \emph{central series} for a group $G$ is a sequence $(G_i)$ of subgroups \[1  = G_0 \lhd G_1 \lhd \cdots \lhd G_n = G\] such that $G_i / G_{i-1} \leq Z(G/G_{i-1})$ for all $i$.}
Note that $G \text{ nilpotent} \iff G$ has a central series.
\dfn{Descending Central Series}{A \emph{descending central series} for a group $G$ is a sequence $(G_i)$ of subgroups \[G = G_0 \triangleright G_1 \triangleright \cdots \triangleright G_n \triangleright \cdots\] where for each $n$, $G_{n+1} = [G_n, G]$. In other words, $L_1(G) = G$, $L_{n+1}(G) = [G, L_n(G)]$.}
Note that $G \text{ nilpotent} \iff G$ has a descending central series which has $n$ such that $L_n(G) = 1$.
\dfn{Ascending Central Series}{An \emph{ascending central series} for a group $G$ is a sequence $(G_i)$ of subgroups \[1 = Z_0 \lhd Z_1 \lhd \cdots \lhd Z_n \lhd \cdots\] where each successive group is defined by $Z_{n+1} = \{g \in G : \forall y \in G, [x, y] \leq Z_i\}$. In other words, $U_0(G) = 1$, $U_{n+1}(G) / U_n(G) = Z(G/U_n(G))$.}
Note that $G \text{ nilpotent} \iff G$ has an ascending central series which has $n$ such that $U_n(G) = G$.

For $G$ nilpotent:

The number of steps it takes the acs to reach $G$ is the same number of steps the dcs takes to reach $1$, which are both the least length of a central series.

\dfn{Nilpotency Degree}{If $G$ nilpotent, $G$ has \emph{nilpotency degree} $t$ if either series takes $t$ steps.}
\ex{Nilpotency Degrees}{
    \begin{itemize}
        \item Degree 0: $G = 1$.
        \item Degree 1: $G$ abelian, $G \neq 1$. 
    \end{itemize}
}
Fact (``normalizer property''):
\begin{itemize}
    \item $G$ nilpotent and $H < G \implies N_G(H) > H$.
\end{itemize}
\begin{proof}
    We induct on nilpotence degree:

    \textbf{Base Case:} $t = 0$ or $t=1$. $G$ abelian. $H < G$. $N_G(H) = G > H$.

    Now suppose the nilpotenecy degree of $G = t +1$ and we have the normalizer property for all groups of nilpotency degree $\leq t$. Let $H < G$. We proceed with case analysis involving the center. We know that $Z(G) \leq N_G(H)$, so:
    \begin{itemize}
        \item Case 1: If $Z(G) \not\leq H$, then $H \neq N_G(H)$, done. 
        \item Case 2: If $Z(G) \leq H$, consider the acs. We have that $U_0 = 1, U_1 = Z(G)$. We take for granted that the nilpotency degree of $G/Z(G) = t$. So consider the acs of $G/Z(G)$, where $U_0 = Z(G)/Z(G) = 1$, $U_1 = G/Z(G)$. So, $Z(G) \leq H < G \implies H/Z(G) < G/Z(G)$ and $G/Z(G)$ has nilpotency degree $t$. So $N_{G/Z(G)}(H/Z(G)) > H/Z(G)$. So $N_G(H) > H$.
    \end{itemize}
    
\end{proof}

\thm{}{If $G_1, G_2$ are niltpotent, then $G_1 \times G_2$ is nilpotent.}
Recall that if $M,N \lhd G$ and $M \cap N = 1$, then $mn = nm$ for all $m \in M$ and $n \in N$, then $MN \cong M \times N$ and $MN \lhd G$. This generalizes for more than 2 subgroups.

Also recall if $G$ finite and $H \in \Syl_p(G)$, then $N_G(H)  = N_G(N_G(H))$.
\begin{proof}
    $H \lhd N_G(H)$, so $H$ is the unique Sylow $p$-subgroup of $N_G(H)$. So now if $g \in N_G(N_G(H))$, then $N_G(H)^g = N_G(H)$, so $H^g$ is the unique Sylow subgroup of $N_G(H)^g = H$, so $g \in N_G(H)$.
\end{proof}

Now let $G$ be finite and nilpotent. Let $|G| = p_1^{t_1}p_2^{t_2}\ldots p_k^{t_k}$. We claim that for each $i$, $G$ has a unique normal Sylow $p_i$-subgroup, $P_i$.
\begin{proof}
    Let $P_i \in \Syl_{p_i}(G)$. Then $N_G(P_i) = N_G(N_G(P_i))$. By the normalizer property, $N_G(P_i) = G$. So $P_i \lhd G$ and $P_i$ is the unique Sylow $p_i$-subgroup of $G$.
\end{proof}
Then by Lagrange, $P_i \cap P_j = 1$ for $i \neq j$. So $P_1P_2\ldots P_k \lhd G$ and $P_1 \times P_2 \times \ldots \times P_k \cong P_1P_2\ldots P_k$. So $P_1P_2 \ldots P_k = G \cong P_1 \times P_2 \times \ldots \times P_k$.

\section{2/7 - Jordan Holder}
\thm{Jordan-Holder}{Let $G$ be a group. Let $(H_i)_{0 \leq i \leq n}$ and $(G_j)_{0 \leq j \leq m}$ be two composition series of $G$. Then $n = m$ and there is a permutation $\sigma$ of $\{0, 1, \ldots, n\}$ such that $H_i / H_{i-1} \cong G_{\sigma(i)} / G_{\sigma(i-1)}$ for all $i$.}
\begin{proof}
    If $G = 1$ or if $G$ is simple, then the result is trivial. So assume $G$ is not simple or trivial so that $m, n > 1$.

    Let $H = H_{n-1}$, so $1 < H \lhd G$. $G/H$ is simple and $(H_i)_{0 \leq i \leq n-1}$ is a composition series for $H$.

    Let $k$ be least $k$ such taht $G_k \not\leq H$. Note that $0 < k < m$. ($G_0 = 1 \leq H, G_m = G \not\leq H$.) Form the subnormal series $(G_iH/H)_{0 \leq j \leq m} \in G/H$. As $G/H$ is simple, this series is a series of $H/H$'s followed by series of $G/H$'s. 

    For $j<k$, $G_j \leq H, G_jH = H \Rightarrow G_jH/H = H/H = 1$.

    So for $j \geq k$, $G_j H/H = G/H$, so $G = G_jH$. 

    For $j \geq k$, $G/H = G_jH/H \cong G_j/(G_j \cap H)$.

    So $(G/H$ simple), $G_j / (G_j \cap H)$ simple for $j \geq k$. 

    Recall $(H_i)_{0 \leq i \leq n-1}$ is for $h$ with $n-1$ steps, so our IH applies to $H$. 

    Define subnormal series in $H$ that is $(G_j \cap H)_{0 \leq j \leq m}$.

    We want to argue two things:
    \begin{enumerate}[label=(\alph*)]
        \item Deleting one repetition in $(G_j \cap H)_{0 \leq j \leq m}$, we obtain a composition series for $H$.
        \item By IH applied to $H$, $m-1 = n-1$ so $m=n$. Argue that we can find IM's between quotients of $G_j$ and $H_j$.
    \end{enumerate}
    Note that for $j < k, G_j \leq H$, so $G_j = G_j \cap H$. 

    Let $j > k$. $G_{j-1} \lhd G_j, G_j \cap H \lhd G_j$. Then we have that $G_{j_1}(G_j \cap H) = G_{j}$. This yields:
    \begin{align*}
        \frac{G_{j-1}(G_j \cap H)}{(G_j \cap H)} &\lhd \frac{G_j}{G_j \cap H} \tag{this RHS is simple} \\
    \end{align*}
    So either $G_{j-1}(G_j \cap H) = G_j$ or $G_{j-1}(G_j \cap H) = G_j \cap H$. However, since $j >k$, we have that $G_{j-1} \not\leq H$, so $G_{j-1} (G_j \cap H) = G_j$. So:
    \begin{align*}
        \frac{G_j}{G_{j-1}} = \frac{G_{j-1}(G_j \cap H)}{G_{j-1}} \\
        \cong \frac{G_j \cap H}{(G_{j} \cap H)\cap G_{j-1}} = \frac{G_j \cap H}{G_{j-1} \cap H}
    \end{align*}
    To finish, we have that $G_{k-1} = G_{k-1} \cap H = G_k \cap H$. We also want to show that $G/H \cong G_k / G_{k-1}$. From $G_{k-1} = G_k \cap H$, $G_k / G_{k-1} \cong G_k / (G_k \cap H) \cong G/H$, by fact already proved. 

    If $G$ has a composition series with $n$ steps, then a $U$ composition series have $n$ steps and ``same quotients''.

    $G \neq 1$, G not simple. So $n \geq 2$. Fix a compostion series $(H_i)_{0 \leq i \leq n}$ so that $H = H_{n-1} \leq G$. 

    Get another composition series $(G_j)_{0 \leq j \leq m}$. Choose $k$ such that $G_k \not\leq H$. Recall that $G_j \cap H = G_j$ for $j < k$ and that for $j>k$, $\frac{G_j \cap H}{G_{j-1} \cap H} \cong \frac{G_j}{G_{j-1}}$.

    We are left to prove that $G_{k-1} = G_k \cap H$ and $\frac{G_k}{G_{k-1}} \cong \frac{G}{H}$. 

    Assessing this, $G_{k-1} \cap H = G_k \cap H$ and seq$(G_j \cap H : j \neq k, 0 \leq j \leq m)$ is a composition series for $H$ with $m-1$ steps with quotients $\frac{G_j}{G_{j-1}}$ for $0 < j  \leq m, j \neq k$. 

    We have 2 composition series for $h$ with 
    \begin{align*}
        (H_i)_{0 \leq i \leq n-1} &\text{ and } (G_j \cap H)_{0 \leq j \leq m-1} \\
    \end{align*}
    By the inductive hypothesis, $m-1 = n-1$, so $m = n$ and up to isomorphism, we can match the quotients of the two series. 
\end{proof}

\section{2/9 - Free Groups}
Recall that $G$ is cyclic if $G = \langle g \rangle$ for some $g \in G$. In this case, there is a unique HM $\phi : (\ZZ, +) \to G$ such that $\phi(1) = g$. Clearly $\phi$ is surjective, so $G \cong \ZZ / \ker \phi$. So $G$ is abelian.

If $\ker(\phi) = 0 = 0 \ZZ$, then $|G| = \infty$, $\phi$ is an isomorphism from $\ZZ$ to $G$. So there is a unique $n > 0$ such that $\ker(\phi) = n \ZZ$ and $|g| = n$. $\phi$ induces an IM from $\ZZ / n \ZZ$ to $G$.

Let's talk about groups that are generated by two elements $G = \langle a, b \rangle$. 
\begin{enumerate}
    \item Any symmetric group $S_n$ for $n \geq 2$ is generated by two elements.
    \begin{align*}
        S_n = \langle (12), (1 \ldots n)\rangle
    \end{align*}
    \item Any dihedral group $D_k$ for finite $k$. 
\end{enumerate}
Our goal is to find a group $F = \langle a, b \rangle$ such that for all $G$ and all $g,h \in G$, there is a unique HM $\phi : F \to G$ such that $\phi(a) = g$ and $\phi(b) = h$. 

We start by trying to construct $F$ using words. Let $\Sigma = \{ a, b\}$. A \emph{word} in $\Sigma$ is a finite sequence of elements of $\Sigma \times \ZZ$. 
\ex{Words}{\begin{itemize}
    \item Write $(a, 3)(b, 5)(a, 2)(b, 1)$ as $a^3b^5a^2b$.
\end{itemize}}
\noindent A word is reduced if
\begin{itemize}
    \item No pairs of form $(s, 0)$ for $s \in \Sigma$
    \item No successive pairs of form $(s, i)(s, j)$ for $s \in \Sigma$.
\end{itemize}
\dfn{Reduction}{Remove terms of the form $s^0$ and replace successive terms of the form $s^is^j$ with $s^{i+j}$.}

\section{2/12 - More Alphabet Stuff}
Let $\sigma$ be a set of any symbols. Words are finite sequences of elements of $\sigma \times \ZZ$. A word is reduced if it has no terms of the form $(s, 0)$ and no successive terms of the form $(s, i)(s, j)$.

Let $W$ be the set of all reduced words in $\sigma^*$. We want to associated to each $s \in \Sigma$, a permutation $\pi_s$ of $W$. The idea is that $\pi_s(w)$ is a unique reduction of $sw$. 

$\pi_s(w)=sw$ makes $w$ start with a term $s^i$ for some $i \in \ZZ$.

If $w = s^i \ldots$ and $i \neq -1$, then $\pi_s(w) = s^{i+1} \ldots$.

If $w = s^{-1}\ldots$, then $\pi_s(w) = \ldots$.

Define $\rho_s$, whose intuition is to reduce $s^{-1}w$. 

$\rho_s(w) = s^{-1}w$ unless $w$ starts with a power of $s$. 

If $w = s^i \ldots$ and $i \neq 1$, then $\rho_s(w) = s^{i-1} \ldots$.

If $w = s \ldots$, then $\rho_s(w) = \ldots$.

One can check that $\pi_s \circ \rho_s = \rho_s \circ \pi_s = \id_W$. As such, $\pi_s$ is a permutation and $\rho_s = \pi_s^{-1}$.

Let $G$ be the subgroup of $\Sigma_w$ (the group of permutations of $w$) generated by $\{\pi_s : s \in \Sigma\}$. If $w$ is a word (not necessarily reduced), say $w = s_1^{i_1}s_2^{i_2}\ldots s_n^{i_n}$, then $\pi_{w} = \pi_{s_1}^{i_1}\pi_{s_2}^{i_2}\ldots \pi_{s_n}^{i_n} \in G$. 

Key fact 1: If $w'$ is obtained from $w$ by a single reduction step, then $\pi_{w'} = \pi_{w}$.

Key fact 2: $G = \{\pi_w : w \in W\}$.
\begin{proof}
    from key fact 1. 
\end{proof}
Key fact 3: If $w \in W$, then $\pi_w(\epsilon) = w$. 

So every $g \in G$ is of form $\pi_w$ for unique reduced word $w$, where $w = g(\epsilon)$. 

Key fact 4: For any words $v, w$, $\pi_v \circ \pi_w = \pi_{vw}$.

Key fact 5: if $w$ is a word and reduced $v$ is obtained from $w$ by some reduction sequence, then $\pi_v(\epsilon) = \pi_w(\epsilon) = v \Rightarrow$ $v$ is determined by $w$. 

Key fact 6: If $v,w \in W$ and $u$ is the unique redcuced word obatined by reducing $vw$, then $\pi_u = \pi_v \circ \pi_w$.

Conclusion: $G$ is isomorphic to the group of whose underlying set is $W$ and whose group operation is ``concatenate and reduce''.

Key point: $G_\Sigma$ has the following ``universal property'': For any group $H$ and any function $\phi : \Sigma \to H$, there is a unique HM $\Phi : G \to H$ such that $\Phi \circ \pi_s = \phi(s)$ for all $s \in \Sigma$.

\dfn{Free Group}{The \emph{free group} on $\Sigma$ is the group $G_\Sigma$ with the universal property.}

\section{2/14 - Free Groups}
Key property of $Free(\Sigma)$ is called the ``universal property.'' Basically, for all groups $H$ and functions $f : \sigma \to H$, there is unique HM $\alpha : Free(\Sigma) \to H$ such that $\alpha(s') = f(s)'$ for all $s \in \Sigma$.

Let's proceed with a proof of the universal property.
\begin{proof}
    \begin{enumerate}
        \item If such an $\alpha$ exists, then it must be given by the formula 
        $$\alpha(s_1^{i_1}s_2^{i_2}\ldots s_n^{i_n}) = f(s_1)^{i_1}f(s_2)^{i_2}\ldots f(s_n)^{i_n}$$
        for all reduced words $s_1^{i_1}s_2^{i_2}\ldots s_n^{i_n}$.
        \item Need to verify that function $\alpha$ specified by $(x)$ is a group HM.
        
        Let $v= s_1^{i_1}s_2^{i_2}\ldots s_n^{i_n}$ and $w = t_1^{j_1}t_2^{j_2}\ldots t_m^{j_m}$ be reduced words. Then $vw$ is a reduced word and $\alpha(vw) = \alpha(v)\alpha(w)$.
    \end{enumerate}
\end{proof}

\section{2/19 - Rings}
What do we mean when we say $R$ is a ring? For now, they will be commutative rings with unity. 

\dfn{Ring Homomorphism}{Let $R$ and $S$ be rings. Then $\phi: R \to S$ is a \emph{ring homomorphism} iff:
\begin{align*}
    \phi(a+b) &= \phi(a) + \phi(b) \\
    \phi(ab) &= \phi(a)\phi(b) \\
    \phi(1_R) &= 1_S
\end{align*}
for all $a,b \in R$.}
\dfn{Subring}{Let $R \subseteq S$ be a subring. This means that the inclusion map $\iota : R \to S$ is a ring homomorphism.}
\dfn{Ideal}{$I \subseteq R$ is an \emph{ideal} iff $I \leq (R, +)$ and $tr \in I$ for all $t \in R$ and $r \in I$. }
\nt{If $\phi: R \to S$ is a ring homomorphism, then $\ker(\phi) = \{r \in R : \phi(r) = 0\}$ is an ideal of $R$.

Also, if $I$ is an ideal, then $I = R$ iff $1 \in I$. 

$\{0\}$ and $R$ are always ideals. `zero ideal'}

\dfn{Quotient Ring}{Let $I \lhd R$. Then $R/I = \{r + I : r \in R\}$ is the \emph{quotient ring} of $R$ by $I$.}
\nt{$(s + I)(t + I) = st + I$. $\phi_I : R \to R/I$ is a quotient homormophism. $\phi_I(r) = r + I$. $\ker(\phi_I) = I$.}

\dfn{$R$-module}{Let $R$ be a ring. $M$ is an \emph{$R$-module} if $M$ is equipped with operations $+: M \times M \to M$ and scalar multiplication$: R \times M \to M$ with the following properties:
\begin{itemize}
    \item $(M, +)$ is an abelian group.
    \item $r(m+n) = rm + rn$ for all $r \in R$ and $m,n \in M$.
    \item $(r+s)m = rm + sm$ for all $r,s \in R$ and $m \in M$.
    \item $(rs)m = r(sm)$ for all $r,s \in R$ and $m \in M$.
    \item $1m = m$ for all $m \in M$.
    \item $0m = 0$ for all $m \in M$.
\end{itemize}}

\ex{}{\begin{enumerate}
    \item If $R$ is a field, $R$-modules are just vector spaces over $R$. 
    \item Let $R = \ZZ$. We claim that $\ZZ$-modules are just abelian groups. Let $G$ be a $\ZZ$-module. Just forget scalar multiplication to get an abelian group.
    
    Converse: Let $(G, +)$ be an abelian group. We have that $ 0g = 0$, $1g = g, (n+1)g = ng + g$, and $(-n)g = 0 - (ng)$. 
    \item Let $K$ be a field, $R = K[x]$. Let $M$ be a $K[x]$-module. 
    \begin{enumerate}
        \item $K$ is a subring of $K[x]$. So $M$ is a $K$-module that is a vector space over $K$.
        \item Let $T: M \to M$ be given by $T(m) = xm$. Then $T$ is a $K$-linear map. Note that we can recover $K[x]$-module structure from $K$-vector space structure and $T$. Namely, $(a_0 + a_1x + \ldots + a_nx^n)m = a_0m + a_1T(m) + \ldots + a_nT^n(m)$.
    \end{enumerate}
    \item For any ring $R$, $R$ is an $R$-module.
    \item Let $M, N$ be $R$-modules. $M \dotplus N = \{(m,n) : m \in M, n \in N\}$ is an $R$-module with coordinatewise operations.
    \item Let $M, N$ be $R$-modules. $Hom_R(M,N) = \{f : M \to N : f \text{ is an $R$-module HM}\}$ is an $R$-module that is $R$ linear. 
\end{enumerate}}
\dfn{Submodule}{If $N$ is an $R$-module then $M \subseteq N$ is a \emph{submodule} iff $M$ is a subgroup of $(N, +)$ and $rm \in M$ for all $r \in R$ and $m \in M$.}
\ex{}{
    \begin{enumerate}
        \item Submodules of vector spaces over $R$ are just subspaces.
        \item In abelian groups, submodules are just subgroups.
        \item If $R$ is a ringe viewed as an $R$-module, then submodules are just ideals.
        \item If $V$ is a vector space over $K$ with $T: V \to V$ linear, view $V$ as a $K[x]$-module. Then submodules of $V$ are just $T$-invariant subspaces.
    \end{enumerate}
}
\dfn{Quotient Module}{If $M \leq N$, $N/M$ is defined in the obvious way. $N/M = \{n + M : n \in N\}$.}
$\phi_M : N \to N/M$ is a quotient homomorphism. 

Also, if $T: M \to N$ is a linear homomorphism, then $\ker(T) = \{m \in M : T(m) = 0\}$. also, $im(T) \cong M / \ker(T)$.
\dfn{Cokernel}{The \emph{cokernel} of $T$ is $N / im(T)$.}
\newpage
\section{Some Random Category Stuff}
% define \span as \operataorname{span}
\def\span{\operatorname{span}}
\dfn{Finitely Generated}{An $R$-module $M$ is \emph{finitely generated} if there is $X \subseteq M$ finite such that $M = \span(X)$.}
\dfn{Noetherian}{An $R$-module $M$ is \emph{Noetherian} if every submodule of $M$ is finitely generated.}
\dfn{Noetherian Ring}{A ring $R$ is \emph{Noetherian} iff $R$ is Noetherian as an $R$-module.}
\ex{}{PID's are Noetherian rings.}
\nt{For any ring $R$, $R$ is the $R$-span of $\{1\}$. So $R$ is finitely generated.}
\ex{Non-Noetherian Ring}{Let $R = \mathbb{Z}[x_1, x_2, \ldots]$ be the ring of polynomials in infinitely many variables. Then $R$ is not Noetherian.

We have that $I = (x_0, x_1, \cdots)$, which is $p \in R$ such that the constant term of $p$ is 0. Supposed by contradiction that $I$ is finitely generated. Say that $I = R$-span of $(p_0, \cdots, p_{n-1})$. That is, $I = (\sum_{i=0}^{n-1} a_i p_i : a_i \in R)$. Then $x_n \in I$, so $x_n = \sum_{i=0}^{n-1} a_i p_i$ for some $a_i \in R$. Let $R$ be so large that $x_k$ does not appear in any $p_j$. $x_k \in I$, so $x_k = \sum_{i=0}^{n-1} a_i p_i$. 

Consider assigning values to $x$ like:
\begin{itemize}
    \item $x_k \to 1$
    \item $x_i \to 0$ for $i \neq k$
\end{itemize}
This gives $1 = \sum_{i=1}^{n-1} a_i  0 = 0$, which is a contradiction.
}
\thm{}{Let $M$ be an $R$-module. Then the following are equivalent:
\begin{enumerate}
    \item $M$ is Noetherian.
    \item Whenever $(N_i)$ such that $N_i \leq N_{i+1}$ and $N_i \leq M$, $N_i$ is eventually constant.
    \item Any nonempty famiy of submodules of $M$ has a maximal element.
\end{enumerate}}
\begin{proof}
    \begin{itemize}
        \item $1 \implies 2$. Let $(N_i)$ be a chain of submodules of $M$. Then $N = \bigcup_{i=1}^\infty N_i$ is a submodule of $M$. $N$ is finitely generated, so fix $X \subseteq N$ where $X$ is finite. $N = \span(X)$, find $i$ such that $X \subseteq N_i$. Now $N_i \subseteq N$, so $N = \span(X) \subseteq N_i$. So $N_i = N_j = N$ for $j \geq i$.
        \item $\neg 1 \implies \neg 2$. 
    \end{itemize}
\end{proof}
\end{document}
